Feature Selection and Classification of \\ Spam on Social Networking Sites

### 2012-05-09 docs/paper/paper.tex
update

This paper investigate several machine learning techniques to 
detect spam in private messages. We train and evaluate Na\"ive Bayes, 
linear regression, and support vector machine (SVM) 
classifiers. Our implementations used a variety of tools, 
including Matlab, ScalaNLP, LIBSVM, Lucene and Spark, 
an in-memory distributed computing framework designed for machine 
learning and iterative computation.

@@

In contrast to research that focuses on dynamically detecting spam based on user 
activity, Irani et al. \cite{irani} show that static features associated with user signups 
on MySpace are enough to train an effective social spam classifier. They note that
C4.5 decision tree algorithms provide better performance than Na\"ive Bayes in this case. 
As in other works, this only examines publicly available profile information collected 
by social honeypots. Private data collected on users including browser features, IP addresses 
and geographic location would conceivably improve classifier performance substantially.

@@

Many of the data mining algorithms used to detect spam and patterns of misuse 
on SNSs are designed with the assumption that the data and the classifier are 
independent. However, in the case of spam, fraud and other malicious content, 
users will often modify their behavior to evade detection, leading to degraded 
classifier performance and the need to re-train classifiers frequently. Several 
researchers tackle this adversarial problem. Dalvi et al. \cite{dalvi} offer a 
modified Na\"ive Bayes classifier to detect and reclassify data taking into account 
the optimal modification strategy that an adversary could choose. Lowd and Meek \cite{lowd} 
provide a framework for reverse engineering a classifier to determine whether 
an adversary can efficiently learn enough about a classifier to effectively defeat it.

\section{Data set}

@@

Our goal in classification was not only to build an effective classifier for spam detection, 
but also to see how the presence of ``expert features", or features based on account and 
message metadata, could affect classifier accuracy. To this end, with of each of the classifiers 
we chose to implement, we evaluated performance on a bag-of-words representation of message data 
as well as a combination of bag-of-words with expert features.

@@

We initially trained and evaluated three classifiers on a small subset of the data: 
Na\"ive Bayes, linear regression and SVM.  We chose an 11k spam and 11k ham sample for 
training, using this particular sample size due to the size of the spam corpus 
section containing recipient age feature data.

\subsection{Na\"ive Bayes}

@@

\subsection{Na\"ive Bayes Results}

The Na\"ive Bayes classifier had an accuracy of 82\% when trained on the bag-of-words. Its accuracy dropped 
to 77\% when expert features were added. The decline in accuracy after adding expert features is most likely 
explained by the ``naive" assumption in this algorithm that the features are independent. Several of 
the features that we used, \eg{account lifetime, and whether a user has friends or photos}, are in 
fact likely to be co-dependent, leading to an overweight of these features during classification.

One of the advantages of the Na\"ive Bayes classifier was that it took very little time to train the classifier, 
and the steps involved were all trivially parallelizable, which makes it convenient for large data sets. 
However, in view of the wealth of expert features available to a SNS, many of which are not independent, this was 
the least compelling classifier choice that we examined.

@@

In this paper, we presented a brief overview of the breakdown of spam and scams on the InterPals web site. 
We then examined a number of statistics on both public and private data that showed that spam users, even when 
grouped together without regard to their specific angle of attack, share a number of characteristics that differentiate 
them from legitimate users. We trained and evaluated Na\"ive Bayes, linear regression and SVM classifiers on a subset 
of the 4 million messages on which we collected feature statistics. Our evaluation compared the performance of 
each classifier on bag-of-word representations of message text vis-\`{a}-vis the combination of bag-of-word features 
with a number of ``expert features." We found that SVMs using an RBF kernel on LIBSVM outperformed the other 
classifiers. In both the linear regression and SVM cases, the addition of expert features substantially increased 
the classifier's accuracy, underscoring the impact of supplementary account-related features in detecting spam on 
SNSs.

>>>>

This paper investigate several machine learning techniques to 
detect spam in private messages. We train and evaluate Naive Bayes, 
linear regression, and support vector machine (SVM) 
classifiers. Our implementations used a variety of tools, 
including Matlab, ScalaNLP, LIBSVM, Lucene and Spark, 
an in-memory distributed computing framework designed for machine 
learning and iterative computation.

@@

In contrast to research that focuses on dynamically detecting spam based on user 
activity, Irani et al. \cite{irani} show that static features associated with user signups 
on MySpace are enough to train an effective social spam classifier. They note that
C4.5 decision tree algorithms provide better performance than Naive Bayes in this case. 
As in other works, this only examines publicly available profile information collected 
by social honeypots. Private data collected on users including browser features, IP addresses 
and geographic location would conceivably improve classifier performance substantially.

@@

Many of the data mining algorithms used to detect spam and patterns of misuse 
on SNSs are designed with the assumption that the data and the classifier are 
independent. However, in the case of spam, fraud and other malicious content, 
users will often modify their behavior to evade detection, leading to degraded 
classifier performance and the need to re-train classifiers frequently. Several 
researchers tackle this adversarial problem. Dalvi et al. \cite{dalvi} offer a 
modified Naive Bayes classifier to detect and reclassify data taking into account 
the optimal modification strategy that an adversary could choose. Lowd and Meek \cite{lowd} 
provide a framework for reverse engineering a classifier to determine whether 
an adversary can efficiently learn enough about a classifier to effectively defeat it.

\section{Data Sets}

@@

Our goal in classification was not only to build an effective classifier for spam detection, 
but also to see how the presence of ``expert features", or features based on account and 
message metadata, could affect classifier accuracy. To this end, with of each of the classifiers 
we chose to implement, we evaluated performance on a bag-of-words representation of message data 
as well as a combination of bag-of-words with expert features. Based on our investigation of 
previous research in this area, we focused our attention on Naive Bayes, linear regression, 
logistic regression and support vector machines, all of which have been used with varying degrees of 
success for email, web and social spam classification, as described in our overview of related work.

@@

We initially trained and evaluated three classifiers on a small subset of the data: 
Naive Bayes, linear regression and SVM.  We chose an 11k spam and 11k ham sample for 
training, using this particular sample size due to the size of the spam corpus 
section containing recipient age feature data.

\subsection{Naive Bayes}

@@

\subsection{Naive Bayes Results}

The Naive Bayes classifier had an accuracy of 82\% when trained on the bag-of-words. Its accuracy dropped 
to 77\% when expert features were added. The decline in accuracy after adding expert features is most likely 
explained by the ``naive" assumption in this algorithm that the features are independent. Several of 
the features that we used, \eg{account lifetime, and whether a user has friends or photos}, are in 
fact likely to be co-dependent, leading to an overweight of these features during classification.

One of the advantages of the Naive Bayes classifier was that it took very little time to train the classifier, 
and the steps involved were all trivially parallelizable, which makes it convenient for large data sets. 
However, in view of the wealth of expert features available to a SNS, many of which are not independent, this was 
the least compelling classifier choice that we examined.

@@

In this paper, we presented a brief overview of the breakdown of spam and scams on the InterPals web site. 
We then examined a number of statistics on both public and private data that showed that spam users, even when 
grouped together without regard to their specific angle of attack, share a number of characteristics that differentiate 
them from legitimate users. We trained and evaluated Naive Bayes, linear regression and SVM classifiers on a subset 
of the 4 million messages on which we collected feature statistics. Our evaluation compared the performance of 
each classifier on bag-of-word representations of message text vis-\`{a}-vis the combination of bag-of-word features 
with a number of ``expert features." We found that SVMs using an RBF kernel on LIBSVM outperformed the other 
classifiers. In both the linear regression and SVM cases, the addition of expert features substantially increased 
the classifier's accuracy, underscoring the impact of supplementary account-related features in detecting spam on 
SNSs.
